---
title: |
    | Credit Risk Measurement
    | Advanced Topics in Banking - Project 2
    | Group 4 | WU Wien, SS 2026
author: "Caterina Piacentini · Farkas Tallos · Balazs Limpek"
date : "2026-05-24"
output:
    pdf_document:
        toc: true
        toc_depth: 3
        number_sections: true
        latex_engine: xelatex
        fig_caption: true
        keep_tex: true
        df_print: kable
        extra_dependencies: ["booktabs", "longtable", "float", "wrapfig", "threeparttable", "threeparttablex", "adjustbox", "array", "multirow", "tikz", "amsmath"]
geometry: "left=2.5cm, right=2.5cm, top=2.5cm, bottom=2.5cm"
fontsize: 11pt
header-includes:
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{float}
  - \floatplacement{figure}{H}
  - \floatplacement{table}{H}
  - \usepackage{xcolor}
  - \definecolor{bawagblue}{HTML}{005493}
  - \definecolor{darkgreen}{HTML}{006400}
  - \definecolor{darkorange}{HTML}{FF8C00}
  - \definecolor{darkred}{HTML}{8B0000}
  - \usepackage{fancyhdr}
  - \pagestyle{fancy}
  - \fancyhf{}
  - \fancyhead[L]{\small BAWAG Group AG --- Risk Analysis 2025}
  - \fancyhead[R]{\small Group 4 | WU Wien 2026}
  - \fancyfoot[C]{\thepage}
  - \renewcommand{\headrulewidth}{0.4pt}
  - \usepackage{caption}
  - \captionsetup[table]{skip=4pt}
  - \AtBeginDocument{\hypersetup{colorlinks=true, linkcolor=bawagblue, urlcolor=bawagblue, citecolor=bawagblue}}
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(
    echo = TRUE, 
    warning = FALSE,
    message = FALSE,
    fig.width = 7,
    fig.height = 5,
    fig.pos = "H",
    out.width = "100%"
)

if(!require("pacman")) install.packages("pacman")

pacman::p_load(
    data.table,
    tidyverse,
    knitr,
    kableExtra,
    ggplot2,
    openxlsx,
    dplyr
)

BASE <- "data/"


ratings <- read.xlsx(paste0(BASE, "Data_Group_4.xlsx"), sheet = "Ratings")
portfolio <- read.xlsx(paste0(BASE, "Data_Group_4.xlsx"), sheet = "Portfolio")


lapply(list(ratings, portfolio), setDT)
lapply(list(ratings, portfolio), str)
lapply(list(ratings, portfolio), head)
lapply(list(ratings, portfolio), summary)

nrow(ratings)
nrow(portfolio)
print(colnames(ratings))
print(colnames(portfolio))

# Correctly parse the dots and format to dashes
colnames(ratings)[-1] <- format(as.Date(colnames(ratings)[-1], format = "%Y.%m.%d"), "%Y-%m-%d")
print(colnames(ratings))

```

# **Task**


##
Based on the rating history of 10,000 representative borrowers, determine the one-year
migration matrix based on the Lando and Skodeberg procedure. In doing so, determine the 
migration matrix using the generator matrix based on at least the first five terms.

Estimation of the generator matrix come from this formula: 
$$\lambda_{ij} = \frac{N_{ij}}{\int Y_{i}(s)ds} \qquad i \neq j \qquad \text{and} \qquad \lambda_{ii} = -\sum_{j \neq i} \lambda_{ij}$$
    
where $N_{ij}$ is the number of observed migrations from rating $i$ to rating $j$ and $Y_{i}(s)$ is the number of borrowers in rating $i$ at time $s$. And the integral: $\int Y_{i}(s)ds$ is the total time spent by all borrowers in rating $i$ during the observation period.
The one-year migration matrix can then be obtained by exponentiating the generator matrix: $W(\Delta t) = e^{\Lambda \cdot \Delta t} = \sum_{k=0}^{\infty} \frac{(\Lambda \cdot \Delta t)^k}{k!}$

```{r, include=TRUE}

ratings_order <- c("AAA", "AA", "A", "BBB", "BB", "B", "CCC/C", "D")
K <- length(ratings_order)

# melt ratings to long format
ratings_long <- melt(
    ratings,
    id.vars = "Obligor",
    variable.name = "date",
    value.name = "rating",
    variable.factor = FALSE
)

ratings_long[, date := as.Date(date)]
ratings_long[, rating := factor(rating, levels = ratings_order)]


setorder(ratings_long, Obligor, date)

nrow(ratings_long)
class(ratings_long$date)
levels(ratings_long$rating)

sum(is.na(ratings_long$rating))
setdiff((unique(as.character(ratings_long$rating))), ratings_order)

head(ratings_long, 10)




ncol(ratings) - 1

head(colnames(ratings), 25)
range(ratings_long$date)
length(unique(ratings_long$date))



# prev rating column
ratings_long[, prev_rating := shift(rating, n = 1L, type = "lag"), by = Obligor]

# drop first rows of each obligot
transitions <- ratings_long[!is.na(prev_rating)]


# cross tabulate: rows = from, cols = to, it means how many times we observed a transition from rating i to rating j
N <- table(from = transitions$prev_rating, to = transitions$rating)

# convert from table to matrix
N <- unclass(N)
storage.mode(N) <- "double"


dim(N)
rownames(N)
colnames(N)

sum(N)

diag(N) / rowSums(N)

N["D",]

N


```

```{r, include=TRUE}

# exact days between ratings
ratings_long[, prev_date := shift(date, n = 1L, type = "lag"), by = Obligor]
ratings_long[, dt := as.numeric((date - prev_date) / 365.25)]

# tapply to apply sum to dt by prev_rating, so we get the total time spent in each rating category
# tapply arguments: 1) the vector to apply the function to, 2) the factor to group by, 3) the function to apply, 4) na.rm = TRUE to ignore NA values
Y <- tapply(ratings_long$dt, ratings_long$prev_rating, sum, na.rm = TRUE)

Y <- Y[ratings_order] # reorder Y according to ratings_order


Y # total time spent in each rating category
length(Y) # number of rating categories
sum(Y) # total time spent in all rating categories
Y / sum(Y) # proportion of time spent in each rating category
# check: total time spent in all ratings should be equal to the total time between the first and last rating for all obligors
cbind(N_rowsum = rowSums(N), Y_times_12 = Y * 12)

# off diagonal elements of the generator matrix, divide the number of transitions by the total time spent in the rating category
Lambda <- sweep(N, MARGIN = 1, STATS = Y, FUN = "/")

diag(Lambda) <- 0

diag(Lambda) <- -rowSums(Lambda)

rowSums(Lambda)
all(abs(rowSums(Lambda)) < 1e-10)

all(Lambda[row(Lambda) != col(Lambda)] >= 0)

all(diag(Lambda) <= 0)

Lambda["D",]

```

```{r, include=TRUE}

expm_taylor <- function(M, k_max) {

    K <- nrow(M)
    P <- diag(K)
    term <- diag(K)
    for (k in 1:k_max) {
        term <- (term %*% M) / k
        P <- P + term
    }
    dimnames(P) <- dimnames(M)
    P
}

P5 <-  expm_taylor(Lambda, k_max = 5)
P20 <- expm_taylor(Lambda, k_max = 20)


rowSums(P5)
rowSums(P20)
max(abs(rowSums(P20) - 1))

range(P5)
range(P20)

P20["D",]

max(abs(P5 - P20))

PD_raw <- P20[, "D"]

PD <- pmax(PD_raw, 0.0005) # set a floor for PD to avoid zero probabilities
data.frame(rating = names(PD), PD_raw, PD)

```

```{r, include=TRUE}

round(P20, 6)

pacman::p_load(expm)
P_expm <- expm::expm(Lambda)
max(abs(P_expm - P20))

PD_raw <- P20[, "D"]
PD <- pmax(PD_raw, 0.0005) # set a floor for

data.frame(
    rating = names(PD),
    PD_raw = round(PD_raw, 6),
    PD = round(PD, 6),
    floored = PD_raw < 0.0005
)





```



# **Task**

## 
Determine the credit risk for the portfolio of 100 borrowers based on the Basel
Standardized Approach. Proceed by using the existing ratings as if they were external ratings.

```{r, include=TRUE}
str(portfolio)
portfolio[, .N, by = Rating][order(-N)]

sum(portfolio$Loan.Volume)
range(portfolio$Loan.Volume)
portfolio[, .N, by = Industry]

rw_table <- data.table(
    Rating = c("AAA", "AA", "A", "BBB", "BB", "B", "CCC/C", "D"),
    RW = c(0.20, 0.20, 0.50, 1.00, 1.00, 1.50, 1.50, 1.50)
)

port_sa <- merge(portfolio, rw_table, by = "Rating", all.x = TRUE)

port_sa[, RWA := Loan.Volume * RW]

head(port_sa)

total_EAD <- sum(port_sa$Loan.Volume)
total_RWA <- sum(port_sa$RWA)
capital_req_SA <- 0.08 * total_RWA

port_sa[, .(
    N = .N,
    EAD = sum(Loan.Volume),
    RW = unique(RW),
    RWA = sum(RWA),
    Capital = 0.08 * sum(RWA)
), by = Rating][order(Rating)]


sum(!is.na(port_sa$RW))

total_RWA / total_EAD

capital_req_SA / total_EAD

```