---
title: |
    | Credit Risk Measurement
    | Advanced Topics in Banking - Project 2
    | Group 4 | WU Wien, SS 2026
author: "Caterina Piacentini · Farkas Tallos · Balazs Limpek"
date : "2026-05-24"
output:
    pdf_document:
        toc: true
        toc_depth: 3
        number_sections: true
        latex_engine: xelatex
        fig_caption: true
        keep_tex: true
        df_print: kable
        extra_dependencies: ["booktabs", "longtable", "float", "wrapfig", "threeparttable", "threeparttablex", "adjustbox", "array", "multirow", "tikz", "amsmath"]
geometry: "left=2.5cm, right=2.5cm, top=2.5cm, bottom=2.5cm"
fontsize: 11pt
header-includes:
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{float}
  - \floatplacement{figure}{H}
  - \floatplacement{table}{H}
  - \usepackage{xcolor}
  - \definecolor{bawagblue}{HTML}{005493}
  - \definecolor{darkgreen}{HTML}{006400}
  - \definecolor{darkorange}{HTML}{FF8C00}
  - \definecolor{darkred}{HTML}{8B0000}
  - \usepackage{fancyhdr}
  - \pagestyle{fancy}
  - \fancyhf{}
  - \fancyhead[L]{\small BAWAG Group AG --- Risk Analysis 2025}
  - \fancyhead[R]{\small Group 4 | WU Wien 2026}
  - \fancyfoot[C]{\thepage}
  - \renewcommand{\headrulewidth}{0.4pt}
  - \usepackage{caption}
  - \captionsetup[table]{skip=4pt}
  - \AtBeginDocument{\hypersetup{colorlinks=true, linkcolor=bawagblue, urlcolor=bawagblue, citecolor=bawagblue}}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
    echo       = TRUE,
    warning    = FALSE,
    message    = FALSE,
    fig.width  = 7,
    fig.height = 5,
    fig.pos    = "H",
    out.width  = "100%"
)

if (!require("pacman")) install.packages("pacman")

pacman::p_load(
    data.table,
    tidyverse,
    knitr,
    kableExtra,
    ggplot2,
    openxlsx,
    scales,
    expm
)

# Portable data path. Defaults to ./data/ relative to this Rmd; override via
# Sys.setenv(ATB_PROJECT_2_BASE = "/some/other/path") before knitting if needed.
BASE      <- Sys.getenv("ATB_PROJECT_2_BASE", unset = "")
DATA_FILE <- if (nzchar(BASE)) {
    file.path(BASE, "data", "Data_Group_4.xlsx")
} else {
    "data/Data_Group_4.xlsx"
}

ratings   <- read.xlsx(DATA_FILE, sheet = "Ratings")
portfolio <- read.xlsx(DATA_FILE, sheet = "Portfolio")

setDT(ratings)
setDT(portfolio)

# Excel date headers may arrive as dotted strings ("2024.12.31") or numeric
# serials ("45657"), depending on the openxlsx version. Handle both.
parse_date_col <- function(x) {
    d <- as.Date(x, format = "%Y.%m.%d")
    if (anyNA(d)) {
        d <- as.Date(suppressWarnings(as.numeric(x)), origin = "1899-12-30")
    }
    d
}
colnames(ratings)[-1] <- format(parse_date_col(colnames(ratings)[-1]), "%Y-%m-%d")
```

# Task 1 --- One-Year Migration Matrix (Lando--Skodeberg)

*Author: Farkas Tallos*

## Case-Study Task

Based on the rating history of 10,000 representative borrowers, determine the
one-year migration matrix based on the Lando and Skodeberg procedure. Determine
the migration matrix using the generator matrix based on at least the first
five terms of the matrix exponential.

## Methodology

The Lando--Skodeberg [2002] estimator models rating dynamics as a
time-homogeneous Markov chain in continuous time. The generator matrix
$\Lambda$ collects the instantaneous transition intensities. Its
maximum-likelihood estimator, given the full rating history of all obligors,
is

$$
\hat\lambda_{ij} \;=\; \frac{N_{ij}}{\int_0^T Y_i(s)\,ds}\,,\quad i \neq j,
\qquad
\hat\lambda_{ii} \;=\; -\sum_{j \neq i} \hat\lambda_{ij},
$$

where $N_{ij}$ is the observed number of migrations from rating $i$ to rating
$j$ during the window $[0, T]$, and $Y_i(s)$ is the number of obligors in
state $i$ at time $s$, so the denominator is the total obligor-time spent in
the from-state (slide p.126 of `01_Slides_QFin_AT in Banking_V1_Handout.pdf`).
The one-year migration matrix is then obtained by exponentiating the
generator,

$$
P(\Delta t) \;=\; e^{\Lambda \cdot \Delta t} \;=\; \sum_{k=0}^{\infty}
\frac{(\Lambda \cdot \Delta t)^k}{k!},
$$

which we evaluate via a truncated Taylor expansion (slide p.132). Default
is modelled as an absorbing state, so the row of $\Lambda$ corresponding to
$\mathrm{D}$ is identically zero.

## State Space and Long-Format Transitions

The 8-state rating universe used in this report is ordered from highest
quality to default, matching the lecture convention:

```{r task1-states-long}
ratings_order <- c("AAA", "AA", "A", "BBB", "BB", "B", "CCC/C", "D")
K <- length(ratings_order)

# Wide -> long: one row per (Obligor, snapshot date, rating).
ratings_long <- melt(
    ratings,
    id.vars         = "Obligor",
    variable.name   = "date",
    value.name      = "rating",
    variable.factor = FALSE
)
ratings_long[, date   := as.Date(date)]
ratings_long[, rating := factor(rating, levels = ratings_order)]
setorder(ratings_long, Obligor, date)

# Add the previous-month rating within each obligor; the first row of each
# obligor receives NA because no prior snapshot exists.
ratings_long[, prev_rating := shift(rating, n = 1L, type = "lag"),
             by = Obligor]
transitions <- ratings_long[!is.na(prev_rating)]
```

```{r task1-preview, echo=FALSE}
kable(
    head(transitions[, .(Obligor, date, prev_rating, rating)], 5),
    booktabs = TRUE,
    caption  = "Sample of obligor transitions (first five rows of the long table)."
) |> kable_styling(latex_options = c("hold_position"))
```

Reshaping from wide to long yields one row per obligor-snapshot observation,
and lagging within `Obligor` produces the `(prev_rating -> rating)` pairs that
feed the migration count. Pairs where `prev_rating` is `NA` (the first
observation of each obligor) carry no transition information and are dropped.
The data span 13 monthly snapshots from 2024-12-31 to 2025-12-31, giving 12
transitions per obligor and a total of $10{,}000 \times 12 = 120{,}000$
transition pairs.

## Migration Counts $N_{ij}$

```{r task1-N}
N <- unclass(table(from = transitions$prev_rating,
                   to   = transitions$rating))
storage.mode(N) <- "double"
```

```{r task1-N-table, echo=FALSE}
kable(
    N,
    booktabs = TRUE,
    caption  = "Migration count matrix $N_{ij}$: observed transitions from row-rating to column-rating.",
    format.args = list(big.mark = ",")
) |> kable_styling(latex_options = c("hold_position", "scale_down"))
```

The count matrix is strongly diagonal-dominant: obligors are far more likely
to stay in their current rating than to migrate. Off-diagonal mass
concentrates immediately adjacent to the diagonal --- single-notch moves
dominate two-notch moves --- consistent with the rating-momentum literature.
The bottom row is identically zero off the diagonal because default is
absorbing in this state space, so no transitions originate from $\mathrm{D}$.
The first few new defaults (4 from A, 10 from BBB, 19 from BB, 65 from B,
161 from CCC/C) are visible in the last column and rise monotonically with
risk, which is the qualitative behaviour the IRB and simulation calibrations
will inherit.

## Time in State $Y_i$

```{r task1-Y}
# Add previous snapshot date alongside previous rating, then compute the
# length of each interval (in years) for the from-state's account.
ratings_long[, prev_date := shift(date, n = 1L, type = "lag"),
             by = Obligor]
ratings_long[, dt := as.numeric(date - prev_date) / 365.25]

Y <- tapply(ratings_long$dt, ratings_long$prev_rating,
            sum, na.rm = TRUE)
Y <- Y[ratings_order]
```

```{r task1-Y-table, echo=FALSE}
kable(
    data.table(Rating = names(Y),
               `Obligor-years` = round(as.numeric(Y), 2)),
    booktabs = TRUE,
    caption  = "Total time spent by all obligors in each rating $i$ (obligor-years)."
) |> kable_styling(latex_options = c("hold_position"))
```

Each obligor contributes the year-fraction between consecutive snapshots to
the rating it occupied at the start of that interval. Summing across
obligors gives the denominator of the Lando--Skodeberg estimator. The
exposure-time distribution mirrors the data's rating distribution: the
investment-grade buckets (BBB, A, BB) accumulate the most obligor-years,
while AAA and CCC/C are sparsely populated, which mechanically inflates the
variance of those rows of $\Lambda$. The total $\sum_i Y_i \approx 9{,}993$
obligor-years almost exactly equals 10,000 obligors $\times$ 1 year of
window; the small shortfall reflects the unequal calendar lengths of
individual months.

## Generator Matrix $\Lambda$

```{r task1-Lambda}
# Off-diagonal entries: lambda_{ij} = N_{ij} / Y_i (sweep divides row i by Y_i).
Lambda <- sweep(N, MARGIN = 1, STATS = Y, FUN = "/")

# Stays are not transitions: wipe the diagonal, then set it to the negative
# row-sum so each row of Lambda sums to zero.
diag(Lambda) <- 0
diag(Lambda) <- -rowSums(Lambda)
```

```{r task1-Lambda-table, echo=FALSE}
kable(
    round(Lambda, 4),
    booktabs = TRUE,
    caption  = "Estimated generator $\\Lambda$ (annualised transition intensities)."
) |> kable_styling(latex_options = c("hold_position", "scale_down"))
```

Off-diagonal entries are instantaneous transition intensities, while the
diagonal equals the negative row-sum so mass is conserved. Row sums are zero
to machine precision (largest absolute deviation
`r format(max(abs(rowSums(Lambda))), scientific = TRUE, digits = 2)`).
Intensities grow sharply as we move down the rating ladder: the magnitude
of the CCC/C diagonal is roughly an order of magnitude larger than the AAA
diagonal, meaning CCC/C-rated obligors leave their bucket much faster than
high-grade names --- which, after exponentiation, will translate into PDs
differing by several orders of magnitude.

## Matrix Exponential --- One-Year Migration Matrix

We implement the truncated Taylor expansion of $e^{\Lambda}$ with the
standard recursion $\mathrm{term}_k = \mathrm{term}_{k-1} \cdot \Lambda / k$,
which avoids any explicit factorial and uses one matrix multiplication per
iteration:

```{r task1-expm}
expm_taylor <- function(M, k_max) {
    K    <- nrow(M)
    P    <- diag(K)            # k = 0 term (identity)
    term <- diag(K)            # running term  M^k / k!
    for (k in seq_len(k_max)) {
        term <- (term %*% M) / k
        P    <- P + term
    }
    dimnames(P) <- dimnames(M)
    P
}

P5  <- expm_taylor(Lambda, k_max = 5)   # case-study minimum
P20 <- expm_taylor(Lambda, k_max = 20)  # essentially exact

trunc_error  <- max(abs(P5 - P20))
expm_compare <- max(abs(P20 - expm::expm(Lambda)))
```

```{r task1-P5-table, echo=FALSE}
kable(
    round(P5 * 100, 3),
    booktabs = TRUE,
    caption  = "One-year migration matrix $P^{(5)}$ from the 5-term Taylor expansion (\\%)."
) |> kable_styling(latex_options = c("hold_position", "scale_down"))
```

```{r task1-P20-table, echo=FALSE}
kable(
    round(P20 * 100, 3),
    booktabs = TRUE,
    caption  = "One-year migration matrix $P^{(20)}$ from the 20-term Taylor expansion (\\%)."
) |> kable_styling(latex_options = c("hold_position", "scale_down"))
```

The case-study requires at least five terms of the series (slide p.132); we
additionally compute the 20-term expansion as a benchmark. The maximum
absolute element-wise difference between $P^{(5)}$ and $P^{(20)}$ is
`r format(trunc_error, scientific = TRUE, digits = 2)`, i.e. the 5-term
truncation is already accurate to about two parts in $10^5$ on every cell.
For external validation, $P^{(20)}$ agrees with `expm::expm(Lambda)` to
`r format(expm_compare, scientific = TRUE, digits = 2)`, which is machine
precision. The diagonal entries of $P^{(20)}$ give the one-year *persistence*
probability for each rating, falling from $\sim$91.6% for AAA to $\sim$61.9%
for CCC/C, and the last row equals $(0, \dots, 0, 1)$ exactly, confirming
that default remains absorbing under the exponential map.

## One-Year PD Vector (with 5 bp Floor)

```{r task1-PD}
PD_raw <- P20[, "D"]
PD     <- pmax(PD_raw, 0.0005)   # 5 basis-point regulatory floor
```

```{r task1-PD-table, echo=FALSE}
PD_tbl <- data.table(
    Rating  = names(PD),
    PD_raw  = round(PD_raw, 6),
    PD      = round(PD,     6),
    Floored = PD_raw < 0.0005
)
kable(
    PD_tbl,
    booktabs = TRUE,
    caption  = "One-year probabilities of default by rating, before and after the 5 bp regulatory floor."
) |> kable_styling(latex_options = c("hold_position"))
```

The raw AAA and AA PDs (approximately
`r format(PD_raw["AAA"], digits = 2, scientific = TRUE)` and
`r format(PD_raw["AA"], digits = 2, scientific = TRUE)`)
lie below the 5 basis-point floor required by the regulator (CRR
Art. 160(1)). This is a well-known feature of investment-grade migration
data: the highest-quality buckets simply do not generate enough defaults in
a finite sample for the empirical estimator to be reliable. The floor is the
supervisor's way of acknowledging that even AAA names carry strictly
positive default risk. From A downward the empirical PD dominates the
floor by at least three orders of magnitude, with CCC/C reaching $\sim$28%.
The full ordering $\text{PD}_{\text{AAA}} < \text{PD}_{\text{AA}} < \cdots <
\text{PD}_{\text{CCC/C}}$ is preserved, which is necessary for the
downstream regulatory and simulation models to behave sensibly.

## Summary --- Inputs for Downstream Tasks

The three objects carried forward into Tasks 3, 4 and 5 are:

- **`PD`** --- 8-vector of one-year PDs by rating, after the 5 bp floor.
  Primary input to the IRB closed-form formula (Task 3) and to both
  Merton-style simulations (Tasks 4 and 5).
- **`P20`** --- 8 $\times$ 8 one-year migration matrix. Useful if a
  downstream task needs the full migration probabilities rather than just
  the default column.
- **`Lambda`** --- the underlying generator. Useful for computing
  migration probabilities at any horizon $\Delta t$ via
  $P(\Delta t) = e^{\Lambda \Delta t}$.

# Task 2 --- Basel Standardized Approach

*Author: Farkas Tallos*

## Case-Study Task

Determine the credit risk for the portfolio of 100 borrowers based on the
Basel Standardized Approach. Proceed by using the existing ratings as if
they were external ratings.

## Regulatory Risk-Weight Lookup (CRR Art. 122)

Under the Standardized Approach the risk weight applied to a corporate
exposure is a step function of the external credit assessment (CRR
Article 122; slide p.208 of
`01_Slides_QFin_AT in Banking_V1_Handout.pdf`). Following the case-study
instruction, we treat the internal ratings as if they were ECAI ratings.
The mapping from the eight-bucket rating universe to risk weights, with
defaulted exposures attracting the 150% weight applicable when specific
provisions cover less than 20% of the outstanding amount, is:

```{r task2-rw-table}
rw_table <- data.table(
    Rating = c("AAA", "AA",  "A",  "BBB", "BB",  "B",  "CCC/C", "D"),
    RW     = c(0.20,  0.20,  0.50, 1.00,  1.00,  1.50, 1.50,    1.50)
)
```

```{r task2-rw-table-render, echo=FALSE}
kable(
    rw_table,
    col.names = c("Rating", "Risk Weight"),
    booktabs  = TRUE,
    caption   = "CRR Art. 122 risk weights for corporate exposures."
) |> kable_styling(latex_options = c("hold_position"))
```

## Per-Obligor Risk-Weighted Assets

```{r task2-rwa}
port_sa <- merge(portfolio, rw_table, by = "Rating", all.x = TRUE)
port_sa[, RWA     := Loan.Volume * RW]
port_sa[, Capital := 0.08 * RWA]
setorder(port_sa, -RWA)
```

```{r task2-top10, echo=FALSE}
kable(
    head(port_sa[, .(Obligor, Rating, Industry, Loan.Volume, RW, RWA, Capital)], 10),
    booktabs    = TRUE,
    caption     = "Ten largest obligors by RWA under the Standardized Approach.",
    format.args = list(big.mark = ",")
) |> kable_styling(latex_options = c("hold_position"))
```

## Aggregated Breakdown by Rating Bucket

```{r task2-by-rating}
by_rating <- port_sa[, .(
    N       = .N,
    EAD     = sum(Loan.Volume),
    RW      = unique(RW),
    RWA     = sum(RWA),
    Capital = sum(Capital)
), by = Rating]

# Order by the canonical rating order, not alphabetical.
by_rating[, Rating := factor(Rating, levels = ratings_order)]
setorder(by_rating, Rating)

# Append totals row.
totals <- data.table(
    Rating  = "Total",
    N       = sum(by_rating$N),
    EAD     = sum(by_rating$EAD),
    RW      = sum(by_rating$RWA) / sum(by_rating$EAD),
    RWA     = sum(by_rating$RWA),
    Capital = sum(by_rating$Capital)
)
by_rating_out <- rbind(
    by_rating[, .(Rating = as.character(Rating), N, EAD, RW, RWA, Capital)],
    totals
)
```

```{r task2-by-rating-render, echo=FALSE}
kable(
    by_rating_out,
    booktabs    = TRUE,
    caption     = "Standardized Approach breakdown by rating bucket (EUR, RW in decimal).",
    format.args = list(big.mark = ","),
    digits      = c(0, 0, 0, 2, 0, 0)
) |> kable_styling(latex_options = c("hold_position"))
```

The portfolio is concentrated in BBB (30 obligors, $\sim$29% of EAD) and
BB (27 obligors, $\sim$27% of EAD), both of which carry the
capital-neutral 100% weight. The eight AA and three AAA names absorb 20%
weights, the seventeen A-rated obligors 50%, while the fifteen
sub-investment-grade names (B and CCC/C combined) are penalised at 150%.
No obligors are in default in the current book, so the SA exposure-weighted
risk weight falls between the 75% (everyone BBB) and 150% (everyone CCC/C)
extremes that would bracket this rating distribution.

## Headline Capital Numbers

```{r task2-headline}
total_EAD      <- sum(port_sa$Loan.Volume)
total_RWA      <- sum(port_sa$RWA)
capital_req_SA <- 0.08 * total_RWA
avg_rw         <- total_RWA / total_EAD
eff_cap_ratio  <- capital_req_SA / total_EAD
```

Total exposure at default is
EUR `r format(total_EAD, big.mark = ",")`,
total RWA is
EUR `r format(total_RWA, big.mark = ",")`,
and the minimum capital requirement (8% of RWA) is
EUR `r format(capital_req_SA, big.mark = ",")`.
The exposure-weighted average risk weight is
`r scales::percent(avg_rw, accuracy = 0.01)`,
and the effective capital-to-EAD ratio is
`r scales::percent(eff_cap_ratio, accuracy = 0.01)`.

## Interpretation and Limitations

The Standardized Approach delivers a minimum capital figure of roughly
EUR 1.17 m, or 7.05% of total exposure. That number is
mechanically driven by the rating distribution and a coarse step-function
of risk weights; it ignores three economically important features of the
portfolio. *First*, it makes no use of the empirical PDs we just estimated
in Task 1: a BBB and a BB obligor receive identical 100% weights despite
their Lando--Skodeberg PDs differing by a factor of more than three.
*Second*, it does not account for concentration risk; the largest single
exposure attracts exactly the same per-euro weight as the smallest.
*Third*, it implicitly assumes independence --- there is no correlation
parameter and therefore no notion of tail or unexpected loss. Tasks 3
(IRB), 4 (2-factor Merton with a common market factor) and 5 (2-factor
Merton with industry factors) progressively relax these assumptions, and
the resulting capital figures are compared in Section \ref{sec:comp}.

# Task 3 --- Basel IRB Approach

*Author: Balazs Limpek*

## Case-Study Task

Determine the credit risk for the portfolio of 100 borrowers based on the
Basel IRB Approach based on the regulatory corporate formula. Use the PDs
from Task 1 (with the 5 bp floor), set the LGDs to 55%, and set the
maturity to 2.5 years.

## Implementation

```{r task3-impl, eval=FALSE}
# IRB corporate formula --- slide p.218 of
# 01_Slides_QFin_AT in Banking_V1_Handout.pdf, CRR Art. 153.
#
# Inputs from the case study brief:
#   PD     <- floored PD vector from Task 1 (object `PD`, minimum 5 bp)
#   LGD    <- 0.55
#   M      <- 2.5
#
# Regulatory correlation as a function of PD:
#   R(PD) = 0.12 * (1 - exp(-50*PD)) / (1 - exp(-50)) +
#           0.24 * (1 - (1 - exp(-50*PD)) / (1 - exp(-50)))
#
# Maturity adjustment:
#   b(PD) = (0.11852 - 0.05478 * log(PD))^2
#
# Per-obligor capital requirement:
#   K = LGD * ( pnorm( (qnorm(PD) + sqrt(R) * qnorm(0.999)) /
#                      sqrt(1 - R) ) - PD ) *
#       (1 + (M - 2.5) * b) / (1 - 1.5 * b)
#
# Risk weight (note the 1.06 scaling factor on slide p.218 --- Basel II
# convention carried through this course):
#   RW  = 1.06 * 12.5 * K
#
# Aggregation:
#   RWA_i     = RW_i * EAD_i
#   Capital_i = K_i  * EAD_i      ( = 8% * RWA_i )

# TODO (Balazs):
# - Join PDs onto portfolio by Rating.
# - Compute R, b, K per obligor (vectorised --- single line each).
# - Aggregate RWA and capital across the portfolio.
# - Report headline numbers in the same format as Section 2.5 above.
```

## Results

*To be completed.*

# Task 4 --- 2-Factor Merton with Common Market Factor

*Author: Balazs Limpek*

## Case-Study Task

Determine the credit risk for the portfolio of 100 borrowers based on a
simulation analysis. The simulation is implemented based on the 2-factor
Merton model with at least 5,000 scenarios. The change in value of each
firm depends on a common market factor and an idiosyncratic factor. Use
the PDs from Task 1 (minimum 5 bp), LGDs of 55%, and weight
$w = 0.37$. Determine the unexpected loss as the 99.9% VaR minus the
expected loss.

## Implementation

```{r task4-impl, eval=FALSE}
# 2-factor Merton with one common market factor --- slide p.223 of
# 01_Slides_QFin_AT in Banking_V1_Handout.pdf.
#
# Inputs from the case study brief:
#   PD       <- floored PD vector from Task 1 (object `PD`)
#   LGD      <- 0.55
#   w        <- 0.37
#   n_sim    <- 5000 minimum; 100,000 recommended for a stable 99.9% quantile
#
# Algorithm per scenario:
#   1. DD_i  = -qnorm(PD_i)
#   2. Draw common market factor   M     ~ N(0, 1)
#   3. Draw idiosyncratic factors  eps_i ~ N(0, 1)   (independent across i)
#   4. Asset value:  V_i = w * M + sqrt(1 - w^2) * eps_i
#   5. Default if V_i < -DD_i          (equivalently V_i < qnorm(PD_i))
#   6. Scenario loss = sum_i ( 1{default_i} * LGD * EAD_i )
#
# Report:
#   EL  = mean(loss)
#   VaR = quantile(loss, 0.999)
#   UL  = VaR - EL                     <-- the case-study deliverable
#
# Use set.seed(...) for reproducibility; agree on a seed with Caterina so
# that Task 4 and Task 5 simulations are directly comparable.

# TODO (Balazs):
# - Build (PD, EAD) vectors of length 100 by joining PDs onto portfolio.
# - Vectorise step 4 with a single matrix of standard normals (100 x n_sim).
# - Produce a loss-distribution histogram and report EL / VaR / UL.
```

## Results

*To be completed.*

# Task 5 --- 2-Factor Merton with Industry Factors

*Author: Caterina Piacentini*

## Case-Study Task

Determine the credit risk for the portfolio of 100 borrowers based on a
simulation analysis with the 2-factor Merton model and at least 5,000
scenarios. The change in value of each firm depends on an industry factor
and an idiosyncratic factor: firms in the same industry exhibit
dependence, firms in different industries do not. There are three market
factors, but only one of them is relevant for each firm. Use the PDs from
Task 1 (minimum 5 bp), LGDs of 55%, and weight $w = 0.46$. Determine
the unexpected loss as the 99.9% VaR minus the expected loss.

## Implementation

```{r task5-impl, eval=FALSE}
# 2-factor Merton with three industry factors --- slide p.224 of
# 01_Slides_QFin_AT in Banking_V1_Handout.pdf, case-study Task 5.
#
# Inputs from the case study brief:
#   PD       <- floored PD vector from Task 1 (object `PD`)
#   LGD      <- 0.55
#   w        <- 0.46
#   n_sim    <- 5000 minimum
#   Industry <- portfolio$Industry   (integer 1, 2, or 3)
#
# Algorithm per scenario:
#   1. DD_i  = -qnorm(PD_i)
#   2. Draw three INDEPENDENT industry factors  F_1, F_2, F_3 ~ N(0, 1)
#   3. Draw idiosyncratic factors                eps_i           ~ N(0, 1)
#   4. Asset value:  V_i = w * F_{Industry_i} + sqrt(1 - w^2) * eps_i
#      (each firm loads on exactly ONE industry factor; firms in different
#       industries are conditionally independent)
#   5. Default if V_i < qnorm(PD_i)
#   6. Scenario loss = sum_i ( 1{default_i} * LGD * EAD_i )
#
# Report:
#   EL  = mean(loss)
#   VaR = quantile(loss, 0.999)
#   UL  = VaR - EL
#
# The 3-industry specification breaks the single common factor of Task 4
# into three uncorrelated drivers. With portfolio diversification across
# industries (42 / 29 / 29 obligors), the UL is expected to be lower than
# Task 4's at the same total weight, even though w is slightly larger
# (0.46 vs 0.37), because the diversification effect dominates.

# TODO (Caterina):
# - Reuse the structure of Task 4 but draw a (3 x n_sim) factor matrix and
#   index into it using portfolio$Industry.
# - Use the same seed as Task 4 for paired comparison.
```

## Results

*To be completed.*

# Comparison and Conclusions {#sec:comp}

*Author: Caterina Piacentini*

## Capital Requirements Across Methods

```{r comparison, eval=FALSE}
# TODO (Caterina): build a single comparison table:
#
#   Method                       Capital (EUR)   % of EAD   Notes
#   --------------------------------------------------------------------------
#   Standardized Approach          1,169,104       7.05%    No PD / no corr
#   IRB (regulatory formula)             ...          ...   Single-factor closed form
#   2-factor Merton (market)             ...          ...   w = 0.37, sim
#   2-factor Merton (industry)           ...          ...   w = 0.46, 3 industries
```

## Discussion

*Placeholder for the cross-method discussion required by the deliverable:*

- *Why the four methods produce different capital numbers (rating-only step
  function vs. continuous PD-dependent regulatory formula vs. simulated tail
  vs. simulated tail with diversification).*
- *Advantages and disadvantages of each method (data requirements,
  computational cost, assumptions, regulatory acceptance).*
- *Which method should BAWAG prefer and why, given a concentrated mid-grade
  corporate book.*

## Limitations and Caveats

- *Single-period horizon and time-homogeneity of the Markov chain.*
- *Constant LGD assumption (55%) ignores downturn-LGD effects.*
- *Sample size for the migration counts (10,000 obligors over 12 months)
  is moderate; CCC/C estimates are noisier than the rest.*
- *Simulation results depend on the random seed and on the number of
  scenarios; report Monte Carlo standard errors for the 99.9% quantile.*
